% Chapter 1

\chapter{Introduction} % Main chapter title

\label{Chapter1} % For referencing the chapter elsewhere, use \ref{Chapter1} 

%----------------------------------------------------------------------------------------

% Define some commands to keep the formatting separated from the content 
\newcommand{\keyword}[1]{\textbf{#1}}
\newcommand{\tabhead}[1]{\textbf{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\file}[1]{\texttt{\bfseries#1}}
\newcommand{\option}[1]{\texttt{\itshape#1}}

%----------------------------------------------------------------------------------------

\section{Motivation/Background}
Combining two concepts into a new concept that combines elements from each original concept is something humans can do rather easily. If we think of blending a horse and a bird, we can easily imagine a winged horse like the mythological Pegasus. Or when blending a car with a boat we can think of an amphibian boat. Formalizing the task of concept blending is not as straightforward. The problem consists of retrieving input data, selecting which parts of the data that should be blended and handling contradictory elements in the final blend.
The goal of this master thesis is to implement an algorithm to provide suggestions of elements from one concept that are suitable for transfer to another. Descriptions of the two concepts are retrieved from a public data source, DBpedia, and NLP tools like WordNet and stanford parser are used to find which elements the concept consists of. Although the concept blending algorithm may be used on any concepts, I will focus mainly on a certain domain such as musical instruments, with the goal of designing new instruments. Further the generated instruments will be evaluated on how useful they are in a musical context. The goal is to find out whether this approach can produce new instruments that make sense and find out how the algorithm can be useful to musicians and instrument makers.

%----------------------------------------------------------------------------------------

\section{Research Questions}
\begin{itemize}
\item What are the characteristics of elements in a concept that are suitable for concept blending?
\item Can a concept blending algorithm generate musical instruments that are reasonable and useful?
\item What are optimal solutions for preparing input data and perform contradiction detection, in the context of using the concept blending algorithm to generate musical instruments?
\end{itemize}
The motivation for these questions is to uncover new knowledge on what the concept blending algorithm can be used for, how it can be used successfully and how it compares to other approaches.


%----------------------------------------------------------------------------------------

\section{Chapter Overview}

%----------------------------------------------------------------------------------------