% Chapter 2

\chapter{Literature review} % Main chapter title

\label{Chapter2} % For referencing the chapter elsewhere, use \ref{Chapter2} 

%----------------------------------------------------------------------------------------

\section{Creativity}
%When looking at different definitions of creativity, it is quite abstract. Here are some definitions of creativity:

%Wikipedia:
%Creativity is a phenomenon whereby something new and somehow valuable is formed. The created item may be intangible (such as an idea, a scientific theory, a musical composition, or a joke) or a physical object (such as an invention, a literary work, or a painting).

%Google/Oxford:
%the use of imagination or original ideas to create something; inventiveness.
%"firms are keen to encourage creativity"
%synonyms:
%inventiveness, imagination, imaginativeness, innovation, innovativeness, originality, individuality; 
%artistry, expressiveness, inspiration, vision, creative power, creative talent, creative gift, creative skill, resourcefulness, ingenuity, enterprise
%"challenging objectives motivate staff and encourage creativity"


Cambridge Dictionary gives the following definition of creativity:
the ability to produce original and unusual ideas, or to make something new or imaginative.

%Dictionary.com
%noun
%1.
%the state or quality of being creative.
%2.
%the ability to transcend traditional ideas, rules, patterns, relationships, or the like, and to create meaningful new ideas, forms, methods, interpretations, etc.; originality, progressiveness, or imagination:
%the need for creativity in modern industry; creativity in the performing arts.
%3.
%the process by which one utilizes creative ability:
%Extensive reading stimulated his creativity.

%Merriam-Webster:
%Definition of creativity
%1:  the quality of being creative
%2:  the ability to create her artistic creativity

Creativity happens in different ways. Sometimes a new invention is found because the right elements were available, and a combination of these created something new. For instance the first electric light was invented when Humphry Davy had access to the world's largest battery at the time, and passing current through a strip of platinum produced light. In other cases creativity is goal-driven. When electric light was commercialized it went through a creative process of choosing the right materials and finding an cost-efficient production method with the goal of making the light-bulb profitable.
%It seems hard to define specifically how creativity is actually done, since it can be done in many ways. 
%The established literature seems to branch into different methods. Examples of creative tasks include invention, problem solving and art. Our approach will follow the theory of conceptual blending.

\parencite{busse1980theories} lists seven categories of creativity theories: psychoanalytic, gestalt, association, perceptual, humanistic, cognitive-developmental and composite theories.
They also draw a line between convergent and divergent problems that the theories are applicable to. The difference is convergent problems have one or few right answers while divergent has many possible solutions. One end is a more goal-driven and problem-solving type of creativity and the other end includes more open-ended activities such as art and conceiving novel ideas.

In the association theories, creativity is thought to come from associations. Arthur Koestler combines associationism and psychoanalytic concepts in what he called the \emph{bisociation} theory \parencite{busse1980theories}. In bisociation, two independent matrices of ideas are combined, guided by subconscious processes. Bisociation inspired the theory of conceptual blending by Gilles Fauconnier and Mark Turner \parencite{fauconnier2002way}.

%----------------------------------------------------------------------------------------

%\section{Computational creativity}


%----------------------------------------------------------------------------------------

\section{Concept blending}
Concept blending is a theory proposed by Gilles Fauconnier and Mark Turner. It works by projecting structure from input mental spaces to a separate, "blended" mental space. \parencite{reference3}
%According to \parencite{reference3}: “concept blending constitutes a cognitive process which allows for the combination of certain elements (and their relations) from originally distinct conceptual spaces into a new unified space combining these previously separate elements and allowing the performance of reasoning and inference over the combination.”

%----------------------------------------------------------------------------------------

\section{Analogy}

%Relasjonar er viktig i analogiar
%Analogiar er viktig for concept blending
%Generalisering?

%Dedre Gentner 1983

%Structure mapping engine is a direct mapping approach

%Generalization based models of analogy
%Smaling 2003
%%HDTP (Schmidt et al. 2014)

%Min løysing er delvis inspirert av analogi ved generalisering?

%----------------------------------------------------------------------------------------

\section{Amalgam}
One approach to concept blending is the use of amalgams. \parencite{Reference4}
An amalgam uses generalization of two concepts to make them compatible for blending through unification of the two generalizations. When parts of the concepts share a similar generalization, but are different, they are not unifiable. But they can unify with the \emph{least general generalization}. To make the concepts unifiable, we can therefore generalize corresponding pairs of parts between the concepts. For instance, when creating an amalgam of a \emph{red French vehicle} and a \emph{German minivan}, we can make them unifiable by generalizing the first to a \emph{red European vehicle}. When unifying we get a \emph{red German minivan}. Or if we generalize the second to a \emph{European minivan}, we can unify to a \emph{red French minivan}.
%An amalgam of two descriptions is a new description that contains parts from these two descriptions. (sitat frå Besold eller Ontanón and Plaza 2010?)

%Subsumption relasjon (hyponym subsuming hypernym)
%All informasjon i hypernymet er sant for hyponymet

%----------------------------------------------------------------------------------------

\section{Meaning and context}
An important part of concept blending is not only representing the parts to blend as words, but also their meaning which depends on the context. \parencite{Reference6} propose that there is a difference between lexical concepts and meaning. Meaning is not a function of language, but arise from the use of language. \parencite{Reference6} provides a theory of lexical concept integration, the Theory of Lexical Concepts and Cognitive Models (LCCM). 

%----------------------------------------------------------------------------------------

%\section{The COINVENT Project}

%----------------------------